Optimal Algorithms for Dynamic Balanced Graph Partitioning

### 2020-05-13 paper/disc20clustering.tex
RM intro

We maintain subsets of frequently communicating nodes as \emph{components}.
Components 
Initially,
each node constitutes a single-node component which we refer to as a \emph{singleton} components, 
and the contained node is an \emph{isolated} node.
We define larger components in terms of smaller components.
Concretely,
given a (sub)sequence of requests,
two components $C_1$ and $C_2$
 \emph{merge} into one component as soon as
 for some pair of nodes $v_1 \in C_1$ and $v_2 \in C_2$,
the frequency of requests $\{v_1,v_2\}$ reaches a certain threshold through the sequence.
We keep all nodes of a~component always collocated in the same cluster,
i.e., when we  move a~node,
we move the whole component that contains it.
A~partition that maintains this invariant is a~\emph{component respecting} partition.
%Intuitively, a component preserves the history of requests between its nodes.

@@

\subsection{Improved Algorithm for Online Rematching} \label{sec:k2}
In this section,
we present \RM,
 an algorithm for the \OBRP restricted to clusters of capacity $k=2$.
We interpret a~pair of nodes collocated in one cluster as a~matched pair.
Hence,
the problem is an online variant of the classic matching problem where
a matched pair can separate from each other in order to ``rematch'' with two other nodes
\maciek{Maybe we use SIDMA's connection to matching instead of this local view. SIDMA: "This can be viewed as an online maximal (re)matching problem: clusters of size two contain (“match”) exactly one pair of nodes, and maximizing pairwise communication within each cluster is equivalent to minimizing inter-cluster communication."}.
This is known as the  \emph{Online Rematching} 
problem and a~(non-strict) 7-competitive algorithm is already given by~\cite{repartition-disc},
in which the ratio comes with an additive factor in $O(\alpha\ell^2)$.
We do not only improve upon their competitive ratio,
but also show that our ratio holds \emph{strictly}
(i.e., with no additive factor).
In addition,
our algorithm is similar
and slightly simpler than the one in~\cite{repartition-disc}, 
while our analysis is significantly more simple and concise,
thanks to the charging scheme we devise here.

>>>>

We maintain subsets of frequently communicating nodes as \emph{components}.
Components 
Initially,
each node constitutes a single-node component which we refer to as a \emph{singleton} component, 
and the contained node is an \emph{isolated} node.
We define larger components in terms of smaller components.
Concretely,
given a (sub)sequence of requests,
two components $C_1$ and $C_2$
 \emph{merge} into one component as soon as
 for some pair of nodes $v_1 \in C_1$ and $v_2 \in C_2$,
the frequency of requests $\{v_1,v_2\}$ reaches a certain threshold through the sequence.
We keep all nodes of a~component always collocated in the same cluster,
i.e., when we  move a~node,
we move the whole component that contains it.
A~partition that maintains this invariant is a~\emph{component respecting} partition.
%Intuitively, a component preserves the history of requests between its nodes.

@@

\subsection{Improved Algorithm for Online Rematching} \label{sec:k2}
In this section,
we present \RM,
 an algorithm for the \OBRP restricted to clusters of capacity $k=2$.
We interpret a~pair of nodes collocated in one cluster as a~``matched'' pair.
Hence,
the problem is an online variant of the maximal matching problem where
%a match is retractable.
%That is, 
a matched pair can separate in order to ``rematch'' with two other nodes.
Rematching is necessary for  maximizing intra-cluster communications,
which is dual to minimizing inter-cluster communications.
This is known as the  \emph{online rematching} 
problem and a~non-strict 7-competitive algorithm is already given by~\cite{repartition-disc},
in which the ratio comes with an additive factor in $O(\alpha\ell^2)$.
We do not only improve upon their competitive ratio,
but also show that our ratio holds \emph{strictly}
(i.e., with no additive factor).
Our algorithm is slightly simpler than the one in~\cite{repartition-disc}, 
while our analysis is significantly  simpler and more concise,
thanks to the charging scheme we devise here.
